---
title: "Genotype analysis"
author: "AGC AK"
date: "2023-02-11"
output: html_document
---

```{r Calling libraries, echo=FALSE, include=FALSE}
library(SingleR)
library(DropletUtils)
library(SingleCellExperiment)
library(Seurat)
library(slingshot)
library(ggplot2)
library(scuttle)
library(celldex)
library(AnnotationHub)
library(org.Hs.eg.db)
library(scater)
library(tidyverse)
library(lubridate)
library(data.table)
library(RColorBrewer)
library(RCurl)
library(randomcoloR)
library(cowplot)
library(viridis)
library(gprofiler2)
library(lme4)



SamplePalette =  c(awr925_CL3_TR1="#000000",
                   awr925_CL3_TR2= "#808080",
                   jem839_CL5_TR2 = "#3399FF", 
                   nuc752_CL4_TR1 = "#FF3300",
                   nuc752_CL4_TR2 = "#CC3300",
                   nuc752_CL6_TR1  = "#990000",
                   SK55_CL9_TR1 = "#FF00FF",
                   SK55_CL9_TR2  = "#CC00CC",
                   SK55_CL49C_TR1   = "#CC66CC",
                   SK301_CL13_TR2  = "#3CB371")

SampleGPalette = c(awr925="#000000", 
                   jem839="#3399FF", 
                   nuc752="#FF3300", 
                   SK55="#FF00FF", 
                   SK301="#3CB371")

nb.cols <- 26
Set26 <- colorRampPalette(brewer.pal(12, "Set3"))(nb.cols)
names(Set26)=1:nb.cols-1

```

```{r load files}

recalc=T
if(recalc){
  load(file = "./output/scRNA_seurat_integrated.Rdata")}
if(!recalc){
  load(file = "./output/scRNA_seurat_integrated.Rdata")
}

```

## Identifying markers LMER
```{r identify markers LMER}

  coldata = seurat_integrated@meta.data
  countdata = as.matrix(seurat_integrated@assays$SCT@scale.data)
  coldata$TR = gsub(".*(TR.)", "\\1", coldata$SampleID)
  
  model = as.formula("x ~ Genotype+(1|TR)+(1|BatchID)")
  
if(recalc){  
  runlmer = function(i, mod, gdata, cdata){
    #cat(i, end="\r")
    cdata$x=gdata[i,]
    res = suppressMessages(lmer(mod, data=cdata, verbose = 0))
    res = summary(res)
    return(res$coefficients[2,])
  }
  
  reslmer <- lapply(1:nrow(countdata), function(i){
    runlmer(i, gdata = countdata, mod = model, cdata = coldata)})
  names(reslmer) = rownames(countdata)
  reslmer <- do.call(rbind, reslmer)
  reslmer <- as.data.frame(reslmer)
  reslmer$p.value = 2*pt(abs(reslmer$`t value`), df=ncol(countdata)-1, lower=FALSE)
  reslmer$adj.p = p.adjust(reslmer$p, method="BH")
}
  

```


```{r,fig.cap="Plot markers LMER"}
ggplot(reslmer, aes(x=Estimate, y=-log10(p.value), col=(adj.p<0.05)))+
  geom_point()
```

## Identifying markers for Genotypes

```{r identify markers for Genotype}
if(recalc){
  Idents(seurat_integrated) = "Genotype"
  
  Genotype.Markers= FindMarkers(seurat_integrated, 
                                assay = "integrated",
                                test.use = "wilcox",
                                ident.1 = "C/C", only.pos = F)
  
  Genotype.Markers <- Genotype.Markers[Genotype.Markers$p_val_adj<0.05,]
  Genotype.Markers %>%
    top_n(n = 20, wt = abs(avg_log2FC)) -> top10Genotype.Markers
  
  Genotype.Markers = Genotype.Markers[order(Genotype.Markers$avg_log2FC),]
}
```


```{r,fig.cap="Heatmap grouped by Genotype"}
DoHeatmap(seurat_integrated, features = rownames(top10Genotype.Markers), 
          group.by = "Genotype", assay = "integrated")

```

## Calculating cell percentages

```{r Calculating the percentages of cell per cluster}
#Calculating the percentages of cell per cluster
number_perCluster<- as.data.frame.matrix(table(Samples=seurat_integrated$SampleID, 
                                               Clusters=scRNA.sce$seurat_clusters))
prop_per_Sample<- prop.table(table(Samples=scRNA.sce$Sample, 
                                   Clusters=scRNA.sce$seurat_clusters), margin = 1)

SampleData = unique(coldata[,c("SampleID", "BatchID", "Genotype", "TR")])


res = apply(prop_per_Sample, 2, function(x){summary(lmer(x~Genotype+(1|TR), data=SampleData))})
ResFreq <- lapply(res, function(x) x$coefficients[2,]) %>% do.call(rbind,.) %>% as.data.frame()

ResFreq$pvalue = 2*pt(abs(ResFreq$`t value`), df=10, lower.tail = F)
ResFreq$p.adj = p.adjust(ResFreq$pvalue)

DT::datatable(ResFreq, extensions = "Buttons",
              filter="top",
              caption = paste0("Effects of Genotype on cluster"),
              options = list(
                pageLength = 10,
                info = FALSE,
                lengthMenu = list(c(15,50, 100, -1),
                                  c("15","50", "100" ,"All")
                ),dom = 'Blfrtip',
                buttons = c('copy', 'csv', 'excel', 'pdf')
              ))
```

## Genewise based cell cluster
```{r genewise by cell cluster, warning=FALSE}

if(recalc){
  reslist.by.cluster=list()
  for(i in unique(seurat_integrated$seurat_clusters)){
    #print(i)
    cellidx=coldata$seurat_clusters==i
    countdata.subset=countdata[,cellidx]
    reslmer.cluster <- lapply(1:nrow(countdata), function(i){
      runlmer(i, gdata = countdata.subset, mod = model, cdata = coldata[cellidx,])})
    reslmer.cluster <- do.call(rbind, reslmer.cluster)
    reslmer.cluster <- as.data.frame(reslmer.cluster)
    reslmer.cluster$p.value = 2*pt(abs(reslmer.cluster$`t value`), df=ncol(countdata)-1, lower=FALSE)
    reslmer.cluster$adj.p = p.adjust(reslmer.cluster$p, method="BH")
    row.names(reslmer.cluster) = rownames(countdata)
    reslist.by.cluster[[i]] = reslmer.cluster
    
  }

  
names(reslist.by.cluster) <- unique(seurat_integrated$seurat_clusters)

reslist.by.cluster = reslist.by.cluster[as.character(sort(unique(seurat_integrated$seurat_clusters)))]

DEX.reslist.by.cluster = lapply(reslist.by.cluster, function(x){rownames(x)[x$adj.p<=0.05 & 
                                                                              x$Estimate>=0.2]})

names(DEX.reslist.by.cluster)  <- names(reslist.by.cluster) 

Gost.reslist.by.cluster = lapply(DEX.reslist.by.cluster, function(x){gost(x, 
                                                                          evcodes = T)})

names(Gost.reslist.by.cluster)  <- names(reslist.by.cluster) 

}



```
## Results Cluster 0
```{r cluster 0}
x="0"
df = reslist.by.cluster[[x]]
df = df[order(df$adj.p, decreasing = F),]
DT::datatable(df, extensions = "Buttons",
              filter="top",
              caption = paste0("Genewise results cluster ",x),
              options = list(
                pageLength = 10,
                info = FALSE,
                lengthMenu = list(c(15,50, 100, -1),
                                  c("15","50", "100" ,"All")
                ),dom = 'Blfrtip',
                buttons = c('copy', 'csv', 'excel', 'pdf')
              )
)

ggplot(df, aes(x=Estimate, y=-log10(p.value), color=(adj.p<0.05)))+
      geom_point()+theme_classic()

gostplot(Gost.reslist.by.cluster[[x]])

```
## Results Cluster 1
```{r cluster 1}
x="1"
df = reslist.by.cluster[[x]]
df = df[order(df$adj.p, decreasing = F),]
DT::datatable(df, extensions = "Buttons",
              filter="top",
              caption = paste0("Genewise results cluster ",x),
              options = list(
                pageLength = 10,
                info = FALSE,
                lengthMenu = list(c(15,50, 100, -1),
                                  c("15","50", "100" ,"All")
                ),dom = 'Blfrtip',
                buttons = c('copy', 'csv', 'excel', 'pdf')
              )
)

ggplot(df, aes(x=Estimate, y=-log10(p.value), color=(adj.p<0.05)))+
      geom_point()+theme_classic()

gostplot(Gost.reslist.by.cluster[[x]])

```
## Results Cluster 2
```{r cluster 2}
x="2"
df = reslist.by.cluster[[x]]
df = df[order(df$adj.p, decreasing = F),]
DT::datatable(df, extensions = "Buttons",
              filter="top",
              caption = paste0("Genewise results cluster ",x),
              options = list(
                pageLength = 10,
                info = FALSE,
                lengthMenu = list(c(15,50, 100, -1),
                                  c("15","50", "100" ,"All")
                ),dom = 'Blfrtip',
                buttons = c('copy', 'csv', 'excel', 'pdf')
              )
)

ggplot(df, aes(x=Estimate, y=-log10(p.value), color=(adj.p<0.05)))+
      geom_point()+theme_classic()

gostplot(Gost.reslist.by.cluster[[x]])

```
## Results Cluster 3
```{r cluster 3}
x="3"
df = reslist.by.cluster[[x]]
df = df[order(df$adj.p, decreasing = F),]
DT::datatable(df, extensions = "Buttons",
              filter="top",
              caption = paste0("Genewise results cluster ",x),
              options = list(
                pageLength = 10,
                info = FALSE,
                lengthMenu = list(c(15,50, 100, -1),
                                  c("15","50", "100" ,"All")
                ),dom = 'Blfrtip',
                buttons = c('copy', 'csv', 'excel', 'pdf')
              )
)

ggplot(df, aes(x=Estimate, y=-log10(p.value), color=(adj.p<0.05)))+
      geom_point()+theme_classic()

gostplot(Gost.reslist.by.cluster[[x]])

```
## Results Cluster 4
```{r cluster 4}
x="4"
df = reslist.by.cluster[[x]]
df = df[order(df$adj.p, decreasing = F),]
DT::datatable(df, extensions = "Buttons",
              filter="top",
              caption = paste0("Genewise results cluster ",x),
              options = list(
                pageLength = 10,
                info = FALSE,
                lengthMenu = list(c(15,50, 100, -1),
                                  c("15","50", "100" ,"All")
                ),dom = 'Blfrtip',
                buttons = c('copy', 'csv', 'excel', 'pdf')
              )
)

ggplot(df, aes(x=Estimate, y=-log10(p.value), color=(adj.p<0.05)))+
      geom_point()+theme_classic()

gostplot(Gost.reslist.by.cluster[[x]])

```
## Results Cluster 5
```{r cluster 5}
x="5"
df = reslist.by.cluster[[x]]
df = df[order(df$adj.p, decreasing = F),]
DT::datatable(df, extensions = "Buttons",
              filter="top",
              caption = paste0("Genewise results cluster ",x),
              options = list(
                pageLength = 10,
                info = FALSE,
                lengthMenu = list(c(15,50, 100, -1),
                                  c("15","50", "100" ,"All")
                ),dom = 'Blfrtip',
                buttons = c('copy', 'csv', 'excel', 'pdf')
              )
)

ggplot(df, aes(x=Estimate, y=-log10(p.value), color=(adj.p<0.05)))+
      geom_point()+theme_classic()

gostplot(Gost.reslist.by.cluster[[x]])

```
## Results Cluster 6
```{r cluster 6}
x="6"
df = reslist.by.cluster[[x]]
df = df[order(df$adj.p, decreasing = F),]
DT::datatable(df, extensions = "Buttons",
              filter="top",
              caption = paste0("Genewise results cluster ",x),
              options = list(
                pageLength = 10,
                info = FALSE,
                lengthMenu = list(c(15,50, 100, -1),
                                  c("15","50", "100" ,"All")
                ),dom = 'Blfrtip',
                buttons = c('copy', 'csv', 'excel', 'pdf')
              )
)

ggplot(df, aes(x=Estimate, y=-log10(p.value), color=(adj.p<0.05)))+
      geom_point()+theme_classic()

gostplot(Gost.reslist.by.cluster[[x]])

```
## Results Cluster 7
```{r cluster 7}
x="7"
df = reslist.by.cluster[[x]]
df = df[order(df$adj.p, decreasing = F),]
DT::datatable(df, extensions = "Buttons",
              filter="top",
              caption = paste0("Genewise results cluster ",x),
              options = list(
                pageLength = 10,
                info = FALSE,
                lengthMenu = list(c(15,50, 100, -1),
                                  c("15","50", "100" ,"All")
                ),dom = 'Blfrtip',
                buttons = c('copy', 'csv', 'excel', 'pdf')
              )
)

ggplot(df, aes(x=Estimate, y=-log10(p.value), color=(adj.p<0.05)))+
      geom_point()+theme_classic()

gostplot(Gost.reslist.by.cluster[[x]])

```
## Results Cluster 8
```{r cluster 8}
x="8"
df = reslist.by.cluster[[x]]
df = df[order(df$adj.p, decreasing = F),]
DT::datatable(df, extensions = "Buttons",
              filter="top",
              caption = paste0("Genewise results cluster ",x),
              options = list(
                pageLength = 10,
                info = FALSE,
                lengthMenu = list(c(15,50, 100, -1),
                                  c("15","50", "100" ,"All")
                ),dom = 'Blfrtip',
                buttons = c('copy', 'csv', 'excel', 'pdf')
              )
)

ggplot(df, aes(x=Estimate, y=-log10(p.value), color=(adj.p<0.05)))+
      geom_point()+theme_classic()

gostplot(Gost.reslist.by.cluster[[x]])

```
## Results Cluster 9
```{r cluster 9}
x="9"
df = reslist.by.cluster[[x]]
df = df[order(df$adj.p, decreasing = F),]
DT::datatable(df, extensions = "Buttons",
              filter="top",
              caption = paste0("Genewise results cluster ",x),
              options = list(
                pageLength = 10,
                info = FALSE,
                lengthMenu = list(c(15,50, 100, -1),
                                  c("15","50", "100" ,"All")
                ),dom = 'Blfrtip',
                buttons = c('copy', 'csv', 'excel', 'pdf')
              )
)

ggplot(df, aes(x=Estimate, y=-log10(p.value), color=(adj.p<0.05)))+
      geom_point()+theme_classic()

gostplot(Gost.reslist.by.cluster[[x]])

```
## Results Cluster 10
```{r cluster 10}
x="10"
df = reslist.by.cluster[[x]]
df = df[order(df$adj.p, decreasing = F),]
DT::datatable(df, extensions = "Buttons",
              filter="top",
              caption = paste0("Genewise results cluster ",x),
              options = list(
                pageLength = 10,
                info = FALSE,
                lengthMenu = list(c(15,50, 100, -1),
                                  c("15","50", "100" ,"All")
                ),dom = 'Blfrtip',
                buttons = c('copy', 'csv', 'excel', 'pdf')
              )
)

ggplot(df, aes(x=Estimate, y=-log10(p.value), color=(adj.p<0.05)))+
      geom_point()+theme_classic()

gostplot(Gost.reslist.by.cluster[[x]])

```
## Results Cluster 11
```{r cluster 11}
x="11"
df = reslist.by.cluster[[x]]
df = df[order(df$adj.p, decreasing = F),]
DT::datatable(df, extensions = "Buttons",
              filter="top",
              caption = paste0("Genewise results cluster ",x),
              options = list(
                pageLength = 10,
                info = FALSE,
                lengthMenu = list(c(15,50, 100, -1),
                                  c("15","50", "100" ,"All")
                ),dom = 'Blfrtip',
                buttons = c('copy', 'csv', 'excel', 'pdf')
              )
)

ggplot(df, aes(x=Estimate, y=-log10(p.value), color=(adj.p<0.05)))+
      geom_point()+theme_classic()

gostplot(Gost.reslist.by.cluster[[x]])

```
## Results Cluster 12
```{r cluster 12}
x="12"
df = reslist.by.cluster[[x]]
df = df[order(df$adj.p, decreasing = F),]
DT::datatable(df, extensions = "Buttons",
              filter="top",
              caption = paste0("Genewise results cluster ",x),
              options = list(
                pageLength = 10,
                info = FALSE,
                lengthMenu = list(c(15,50, 100, -1),
                                  c("15","50", "100" ,"All")
                ),dom = 'Blfrtip',
                buttons = c('copy', 'csv', 'excel', 'pdf')
              )
)

ggplot(df, aes(x=Estimate, y=-log10(p.value), color=(adj.p<0.05)))+
      geom_point()+theme_classic()

gostplot(Gost.reslist.by.cluster[[x]])

```
## Results Cluster 13
```{r cluster 13}
x="13"
df = reslist.by.cluster[[x]]
df = df[order(df$adj.p, decreasing = F),]
DT::datatable(df, extensions = "Buttons",
              filter="top",
              caption = paste0("Genewise results cluster ",x),
              options = list(
                pageLength = 10,
                info = FALSE,
                lengthMenu = list(c(15,50, 100, -1),
                                  c("15","50", "100" ,"All")
                ),dom = 'Blfrtip',
                buttons = c('copy', 'csv', 'excel', 'pdf')
              )
)

ggplot(df, aes(x=Estimate, y=-log10(p.value), color=(adj.p<0.05)))+
      geom_point()+theme_classic()

gostplot(Gost.reslist.by.cluster[[x]])

```
## Results Cluster 14
```{r cluster 14}
x="14"
df = reslist.by.cluster[[x]]
df = df[order(df$adj.p, decreasing = F),]
DT::datatable(df, extensions = "Buttons",
              filter="top",
              caption = paste0("Genewise results cluster ",x),
              options = list(
                pageLength = 10,
                info = FALSE,
                lengthMenu = list(c(15,50, 100, -1),
                                  c("15","50", "100" ,"All")
                ),dom = 'Blfrtip',
                buttons = c('copy', 'csv', 'excel', 'pdf')
              )
)

ggplot(df, aes(x=Estimate, y=-log10(p.value), color=(adj.p<0.05)))+
      geom_point()+theme_classic()

gostplot(Gost.reslist.by.cluster[[x]])

```



```{r save}
save(reslist.by.cluster, DEX.reslist.by.cluster, top10Genotype.Markers, Gost.reslist.by.cluster, ResFreq, Genotype.Markers, reslmer, seurat_integrated, scRNA.sce, Cellpalette, SampleGPalette, SamplePalette, prop_per_Sample, file = "./output/scRNA_seurat_integrated.Rdata")

```


